\section{Operational Intelligence Suite}

The platform provides five interconnected intelligence modules that transform raw operational data into actionable insights.

\subsection{Review Intelligence}

Multi-platform scraping (Yelp, Google) with anti-bot detection feeds reviews into ByteDance Seed 1.6 for sentiment analysis. The LLM extracts \textbf{5-category opinions} (food, service, atmosphere, value, cleanliness) and flags reviews requiring manager attention. Batch processing (25 reviews/batch) enables efficient categorization at scale.

\subsection{Menu Intelligence}

Items are ranked by a composite score combining demand and profitability:
\begin{equation}
\text{score} = (\text{orders/day})_{\text{norm}} \times 0.5 + \frac{\text{price} - \text{cost}}{\text{price}} \times 0.5
\end{equation}

The system generates \textbf{pricing recommendations} via a demand-margin matrix (high demand + low margin $\rightarrow$ +10-15\% price) and flags underperformers (score $<25$) for ``86'' removal with auto-generated reasoning.

\subsection{Server Intelligence}

Waiter performance uses Z-score normalization across three weighted metrics:
\begin{equation}
\text{performance} = 0.3 \cdot z_{\text{turn\_time}}^{-1} + 0.4 \cdot z_{\text{tip\_pct}} + 0.3 \cdot z_{\text{covers}}
\end{equation}

Percentile-based tier assignment (strong $\geq$ p75, standard p25--p75, developing $<$ p25) drives routing priority. LLM generates personalized insights: strengths, areas to watch, and improvement suggestions.

\subsection{Scheduling Intelligence}

\textbf{Demand forecasting} applies exponential decay ($0.85^{\text{weeks\_ago}}$) to historical data with linear trend detection. The score-and-rank algorithm optimizes assignments:
\begin{equation}
\text{score} = 0.5 \cdot \text{constraints} + 0.3 \cdot (\text{fairness\_impact} + 50) + 0.2 \cdot \text{preferences}
\end{equation}

Hard constraints (availability, max hours, no overlaps) exclude candidates; soft constraints (shift preferences, avoid clopening) are scored. \textbf{Gini coefficient} targets $<0.25$ for equitable hours distribution. Each assignment includes LLM-generated reasoning.

\subsection{Analytics Engine}

Pre-computed metrics power real-time dashboards:

\begin{table}[h]
\centering\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Metric} & \textbf{Granularity} & \textbf{Application} \\
\midrule
Table Turn Time & Per-visit, waiter, restaurant & Efficiency benchmarking \\
Wait Time & Hourly/daily/weekly & Capacity planning \\
Peak Hours & Day-of-week patterns & Staffing optimization \\
Revenue Trends & Monthly rollups & Financial forecasting \\
Forecast Accuracy & MAPE over 8--12 weeks & Model validation \\
\bottomrule
\end{tabular}
\end{table}

All analytics endpoints support both aggregate views and drill-down to individual staff, shifts, or time periods.
